---
title: "HW4"
subtitle: "Statistical Method"
date: today
author: H24101222_陳凱騫
format:
 pdf:
    include-in-header:
      - text: |
         \usepackage{setspace,relsize}
         \usepackage{geometry}
         \geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
         \setmonofont{Microsoft JhengHei UI}  
         \usepackage{hyperref}
mainfont: "Microsoft JhengHei UI"
# mainfont: "Microsoft JhengHei Bold"
toc: true
documentclass: article
pdf-engine: xelatex
execute:
  tidy: true
  echo: true
  warning: false
  message: false
---

# Load datasets Set1.csv and Set2.csv

```{r}
#| echo: true
#| code-fold: show
library(knitr)
data1 <- read.csv("Set1.csv")
head(data1)
data2 <- read.csv("Set2.csv")
head(data2)
```

# EDA

```{r}
#| output: asis
#| code-fold: false
#| echo: false
library(Hmisc)
latex(describe(data1), file = "", caption.placement = "top")
latex(describe(data2), file = "", caption.placement = "top")
```

```{r}
boxplot(c(data1),horizontal=TRUE,main="boxplot of Set1")
boxplot(c(data2),horizontal=TRUE,main="boxplot of Set2")

```

# (1.a) Use the method of moments (MME) to estimate the model parameters.

## Normal

-   The decision variables for the normal distribution are $\mu$ and $\sigma^{2}$.
-   Objective function: $E(X)=\mu$ and $E(X^{2})= \sigma^{2} + \mu^{2}$
-   Theoretical derivation / efficient algorithms: $$
      \hat{\mu}_{MME}=\frac{1}{n}\sum_{i=1}^{n}x_{i}
      $$ and $$
      \hat{\sigma^{2}_{MME}}=\frac{1}{n}\sum_{i=1}^{n}x_{i}^{2}-\hat{\mu}_{MME}^{2}
      $$

### Set1

```{r}
n <- nrow(data1) # 樣本數量
EX <- mean(data1$x) # 一次動差
EX2 <- mean(data1$x^2) # 二次動差
mu.mme <- EX
sigma2.mme <- EX2 - (EX)^2
print(c(mu.mme,sigma2.mme))
```

```{r}
### Verification: By fitting plot
hist(data1$x, probability = TRUE)
x0 <- seq(-2, 3, 0.01)
lines(x0, dnorm(x0, mu.mme, sqrt(sigma2.mme)), col = 2)
plot(ecdf(data2$x))
lines(x0, pnorm(x0, mu.mme, sqrt(sigma2.mme)), col = 2)
```

### Set2

```{r}
n <- nrow(data2) # 樣本數量
EX <- mean(data2$x) # 一次動差
EX2 <- mean(data2$x^2) # 二次動差
mu.mme <- EX
sigma2.mme <- EX2 - (EX)^2
print(c(mu.mme,sigma2.mme))
```

```{r}
### Verification: By fitting plot
hist(data2$x, probability = TRUE)
x0 <- seq(8.5, 32, 0.01)
lines(x0, dnorm(x0, mu.mme, sqrt(sigma2.mme)), col = 2)
plot(ecdf(data2$x))
lines(x0, pnorm(x0, mu.mme, sqrt(sigma2.mme)), col = 2)
```

## Exponential distribution

-   The decision variable for the exponential distribution is $\lambda$.
-   Objective function: $E(X)=\frac{1}{\lambda}$
-   Theoretical derivation / efficient algorithms: 
  $$
  \hat{\lambda}_{MME}=\frac{n}{\sum_{i=1}^{n}x_{i}}
  $$

### Set1

```{r}
n <- nrow(data1) # 樣本數量
EX <- mean(data1$x) # 一次動差
lambda.mme <- 1/EX
cat("lambda.mme是:")
print(lambda.mme)
```

```{r}
### Verification: By fitting plot
hist(data1$x, probability = TRUE)
x0 <- seq(min(data1$x), max(data1$x), 0.01)
lines(x0, dexp(x = x0, rate = lambda.mme), col = 2)
plot(ecdf(data1$x))
lines(x0, pexp(x0, rate = lambda.mme), col = 2)
```

### Set2

```{r}
n <- nrow(data2) # 樣本數量
EX <- mean(data2$x) # 一次動差
lambda.mme <- 1/EX
cat("lambda.mme是:")
print(lambda.mme)
```

```{r}
### Verification: By fitting plot
hist(data2$x, probability = TRUE)
x0 <- seq(min(data2$x), max(data2$x), 0.01)
lines(x0, dexp(x = x0, rate = lambda.mme), col = 2)
plot(ecdf(data2$x))
lines(x0, pexp(x0, rate = lambda.mme), col = 2)
```

## Laplace distribution

-   The decision variables for the normal distribution are $\mu$ and $b$.

-   Objective function: $E(X)=\mu$ and $E(X^2) = \sigma^2 + \mu^2$

-   Theoretical derivation / efficient algorithms: 
  $$
  \hat{\mu}_{MME}=\frac{\sum^{n}_{i=1}x_{i}}{n}
  $$
  $$
  \hat{b}_{MME} = \sqrt{\frac{1}{2} \left( \frac{1}{n} \sum_{i=1}^{n} X_i^2 - \left( \frac{1}{n} \sum_{i=1}^{n} X_i \right)^2 \right)} 
  $$

### Set1

```{r}
n <- nrow(data1) # 樣本數量
EX <- mean(data1$x) # 一次動差
EX2 <- mean(data1$x^2) # 二次動差
mu.mme <- EX
b.mme <- sqrt(0.5*(EX2-EX^2))
cat("mu.mme是:")
print(mu.mme)
cat("b.mme是:")
print(b.mme)
```

```{r}
### Verification: By fitting plot
# install.packages("extraDistr")
library(extraDistr)
hist(data1$x, probability = TRUE)
x0 <- seq(min(data1$x), max(data1$x), 0.001)
lines(x0, dlaplace(x0, mu = mu.mme, sigma= b.mme), col = 2, lwd =1.2)
plot(ecdf(data1$x))
lines(x0, plaplace(x0, mu = mu.mme, sigma = b.mme), col = 2, lwd=1.5)
```

### Set2

```{r}
n <- nrow(data2) # 樣本數量
EX <- mean(data2$x) # 一次動差
EX2 <- mean(data2$x^2) # 二次動差
mu.mme <- EX
b.mme <- sqrt(0.5*(EX2-EX^2))
cat("mu.mme是:")
print(mu.mme)
cat("b.mme是:")
print(b.mme)
```

```{r}
### Verification: By fitting plot
hist(data2$x, probability = TRUE)
x0 <- seq(min(data2$x), max(data2$x), 0.001)
lines(x0, dlaplace(x0, mu = mu.mme, sigma= b.mme), col = 2, lwd =1.2)
plot(ecdf(data2$x))
lines(x0, plaplace(x0, mu = mu.mme, sigma = b.mme), col = 2, lwd=1.5)
```

## Gamma distribution

-   The decision variables for the normal distribution are $\alpha$ and $\lambda$.

-   Objective function: $E(X)=\frac{\alpha}{\lambda}$ and $E(X^2) = \sigma^2 + EX^2= \frac{\alpha(\alpha+1)}{\lambda^2}$

-   Theoretical derivation / efficient algorithms: 
  $$
  \hat{\alpha}_{MME}=\frac{E(X)^2}{E(X^2)-E(X)^2}
  $$

  $$
  \hat{\lambda}_{MME} = \frac{E(X)}{E(X^2)-E(X)^2}
  $$

### Set1

```{r}
n <- nrow(data1) # 樣本數量
EX <- mean(data1$x) # 一次動差
EX2 <- mean(data1$x^2) # 二次動差
alpha.mme <- EX^2/(EX2-(EX^2))
lambda.mme <- EX/(EX2-(EX^2))
cat("alpha.mme是:")
print(alpha.mme)
cat("lambda.mme是:")
print(lambda.mme)
```

```{r}
### Verification: By fitting plot
# install.packages("extraDistr")
library(extraDistr)
hist(data1$x, probability = TRUE)
x0 <- seq(min(data1$x), max(data1$x), 0.001)
lines(x0, dgamma(x0, shape = alpha.mme, rate = lambda.mme), col = 2, lwd =1.2)
plot(ecdf(data1$x))
lines(x0, pgamma(x0, shape = alpha.mme, rate = lambda.mme), col = 2, lwd=1.5)
```

### Set2

```{r}
n <- nrow(data2) # 樣本數量
EX <- mean(data2$x) # 一次動差
EX2 <- mean(data2$x^2) # 二次動差
alpha.mme <- EX^2/(EX2-(EX^2))
lambda.mme <- EX/(EX2-(EX^2))
cat("alpha.mme是:")
print(alpha.mme)
cat("lambda.mme是:")
print(lambda.mme)
```

```{r}
### Verification: By fitting plot
# install.packages("extraDistr")
library(extraDistr)
hist(data2$x, probability = TRUE)
x0 <- seq(min(data2$x), max(data2$x), 0.001)
lines(x0, dgamma(x0, shape = alpha.mme, rate = lambda.mme), col = 2, lwd =1.2)
plot(ecdf(data2$x))
lines(x0, pgamma(x0, shape = alpha.mme, rate = lambda.mme), col = 2, lwd=1.5)
```

## Conclusion

| Distribution | Dataset1 | Dataset2 |
|--------------|----------|----------|
| Normal       | ($\hat{\mu}$, $\hat{\sigma^2}$)= (0.025, 0.707) | ($\hat{\mu}$, $\hat{\sigma^2}$)= (19.247, 25.913)         |
| Exponential  | $\hat{\lambda}$=40.468                          | $\hat{\lambda}$=0.052        |
| Laplace      | ($\hat{\mu}$, $\hat{b^2}$)=(0.025,0.594)                               | ($\hat{\mu}$, $\hat{b^2}$)=(19.247,3.599)         |
| Gamma        | ($\hat{\alpha}$,$\hat{\lambda}$)=(0.00086,0.03496)                               | ($\hat{\alpha}$,$\hat{\lambda}$)=(14.296,0.743)         |

: MME to estimate the model parameters

# (1.b) Use the maximum likelihood estimates (MLEs) to estimate the model parameters.

## Normal

-   The decision variables for the normal distribution are $\mu$ and $\sigma^{2}$.
-   Objective function: $max_{(\mu,\sigma^{2})}\prod_{i=1}^{n}f(x_i;\mu,\sigma^2)$
-   Theoretical derivation / efficient algorithms: 
  $$
  \hat{\mu}_{MLE}=\frac{1}{n}\sum_{i=1}^{n}x_{i}
  $$ 
  and 
  $$
  \hat{\sigma^{2}_{MlE}}=\frac{1}{n}\sum_{i=1}^{n}(x_{i}-\bar{x})^{2}
  $$

### Set1

```{r}
n <- length(data1$x)
mu.mle <- mean(data1$x)
sig2.mle <- sum((data1$x - mu.mle)^2)/n

### Numerical method ###
likelihood.normal <- function(par, data){
  mu <- par[1]
  sig2 <- par[2]
  joint <- dnorm(data, mean = mu, sd = sqrt(sig2))
  #return(prod(joint))
  #return(sum(log(joint)))
  return(-sum(log(joint)))
}

likelihood.normal(c(1, 2), data = data1$x) # -loglikelihood
likelihood.normal(c(mu.mle, sig2.mle), data = data1$x) # -loglikelihood
#optim(c(2, 1), likelihood.normal, data = dat.x, control=list(fnscale=-1))
mle.norm <- optim(c(2, 1), likelihood.normal, data = data1$x, hessian = TRUE)

```

```{r}
### Verification: By hessian matrix
mle.norm
det(mle.norm$hessian)
```

### Set2

```{r}
### Theoritical method ###
n <- length(data2$x) 
mu.mle <- mean(data2$x)
sig2.mle <- sum((data2$x - mu.mle)^2)/n

### Numerical method ###
likelihood.normal <- function(par, data){
  mu <- par[1]
  sig2 <- par[2]
  joint <- dnorm(data, mean = mu, sd = sqrt(sig2))
  #return(prod(joint))
  #return(sum(log(joint)))
  return(-sum(log(joint)))
}

likelihood.normal(c(1, 2), data = data2$x) # -loglikelihood
likelihood.normal(c(mu.mle, sig2.mle), data = data2$x) # -loglikelihood
  #optim(c(2, 1), likelihood.normal, data = dat.x, control=list(fnscale=-1))
mle.norm <- optim(c(mu.mle, sig2.mle), likelihood.normal, data = data2$x, hessian = TRUE)

```

```{r}
### Verification: By hessian matrix
mle.norm
det(mle.norm$hessian)
```

## Exponential

-   The decision variable for the exponential distribution is $\lambda$.
-   Objective function: $\prod^{n}_{i=1}\lambda e^{\lambda x_{i}}$
-   Theoretical derivation / efficient algorithms: 
$$
\hat{\lambda}=\frac{n}{\sum^{n}_{i=1}x_{i}}
$$

### Set1

```{r}
#| error: true
### Theoritical method ###
n <- length(data1$x)
lambda.mle <- 1/mean(data1$x)
lambda.mle
### Numerical method ###
likelihood.exp <- function(par, data){
  lambda <- par[1]
  joint <- dexp(data,rate = par)
  #return(prod(joint))
  #return(sum(log(joint)))
  return(-sum(log(joint)))
}

likelihood.exp(c(lambda.mle), data = data1$x) # -loglikelihood
#optim(c(2, 1), likelihood.normal, data = dat.x, control=list(fnscale=-1))
mle.exp <- optim(c(40), likelihood.exp, data = data1$x)

```
從EDA中可以得知資料有負值，因此無法在此分布做MLE。

### Set2

```{r}
### Theoritical method ###
n <- length(data2$x) 
lambda.mle <- 1/mean(data2$x)
lambda.mle
### Numerical method ###
likelihood.exp <- function(par, data){
  lambda <- par[1]
  joint <- dexp(data,rate = par)
  #return(prod(joint))
  #return(sum(log(joint)))
  return(-sum(log(joint)))
}
likelihood.exp(c(1), data = data2$x) # -loglikelihood
likelihood.exp(c(lambda.mle), data = data2$x) # -loglikelihood
#optim(c(2, 1), likelihood.normal, data = dat.x, control=list(fnscale=-1))
mle.exp <- optim(1, likelihood.exp, data = data2$x, method = "Brent", lower = 1e-6, upper = 10,hessian = TRUE)

```

```{r}
### Verification: By hessian matrix
mle.exp
```

## Laplace

-   The decision variables for the normal distribution are $\mu$ and $b$.

-   Objective function: 
$$
\prod^{n}_{i=1}\frac{1}{2b}e^{-\frac{|x_{i}-\mu|}{b}}
$$

-   Theoretical derivation / efficient algorithms: 
$$
\hat{\mu} = \text{median}(\{x_i\})
$$
$$
\hat{b} = \frac{1}{n} \sum_{i=1}^n |x_i - \hat{\mu}|
$$

### Set1

```{r}
### Theoritical method ###
n <- length(data1$x)
mu.mle <- median(data1$x)  # Median for mu
b.mle <- mean(abs(data1$x - mu.mle))  # Mean absolute deviation for b
### Numerical method ###
# Negative log-likelihood function for Laplace distribution
library(extraDistr)
likelihood.laplace <- function(par, data) {
  mu <- par[1]
  b <- par[2]
  if (b <= 0) return(Inf)  # Ensure valid parameter range
  -sum(dlaplace(data, mu = mu, sigma = b, log = TRUE))  # Negative log-likelihood
}
init <- c(mu.mle, b.mle)
# Perform optimization
mle.laplace <- optim(
  par = init,
  fn = likelihood.laplace,
  data = data1$x,
  method = "L-BFGS-B",
  lower = c(-Inf, 1e-6),  # Lower bounds
  upper = c(Inf, Inf),    # Upper bounds
  hessian = TRUE          # Compute Hessian matrix
)
# Extract results
mle.laplace
```

### Set2

```{r}
### Theoritical method ###
n <- length(data2$x) 
mu.mle <- median(data2$x)  # Median for mu
b.mle <- mean(abs(data2$x - mu.mle))  # Mean absolute deviation for b
### Numerical method ###
# Negative log-likelihood function for Laplace distribution
library(extraDistr)
likelihood.laplace <- function(par, data) {
  mu <- par[1]
  b <- par[2]
  if (b <= 0) return(Inf)  # Ensure valid parameter range
  -sum(dlaplace(data, mu = mu, sigma = b, log = TRUE))  # Negative log-likelihood
}
init <- c(mu.mle, b.mle)
# Perform optimization
mle.laplace <- optim(
  par = init,
  fn = likelihood.laplace,
  data = data2$x,
  method = "L-BFGS-B",
  lower = c(-Inf, 1e-6),  # Lower bounds
  upper = c(Inf, Inf),    # Upper bounds
  hessian = TRUE          # Compute Hessian matrix
)
# Extract results
mle.laplace
```

## Gamma

- The decision variables for the normal distribution are $\alpha$ and $\lambda$.
- Objective function: 
$$
\prod^{n}_{i=1}\frac{\lambda^\alpha x_{i}^{\alpha - 1} e^{-\lambda x_{i}}}{\Gamma(\alpha)}
$$
- Gamma的參數估計(MLE)不好寫成close form 

### Set1

```{r}
#| error: true
### Numerical method ###
# Negative log-likelihood function for Gamma distribution (rate parameterization)
library(extraDistr)
likelihood.gamma <- function(par, data) {
  alpha <- par[1]
  lambda <- par[2]
  
  if (alpha <= 0 || lambda <= 0) return(Inf)  # Ensure valid parameter range
  # Negative log-likelihood (using dgamma from extraDistr package)
  -sum(dgamma(data, shape = alpha, rate = lambda, log = TRUE))  # Negative log-likelihood
}

# Initial guesses for parameters (alpha, lambda)
alpha.mle <- 2   # Example initial guess for alpha
lambda.mle <- 1  # Example initial guess for lambda
init <- c(alpha.mle, lambda.mle)
# Perform optimization
mle.gamma <- optim(
  par = init,
  fn = likelihood.gamma,
  data = data1$x,
  method = "L-BFGS-B",
  lower = c(0.1, 1e-6),  # Lower bounds for alpha and lambda
  upper = c(Inf, Inf),   # Upper bounds for alpha and lambda
  hessian = TRUE         # Compute Hessian matrix for parameter uncertainty
)

# Extract results
mle.gamma
```
因為data有負號，因此gamma不適合

### Set2

```{r}
### Numerical method ###
# Negative log-likelihood function for Gamma distribution (rate parameterization)
library(extraDistr)
likelihood.gamma <- function(par, data) {
  alpha <- par[1]
  lambda <- par[2]
  if (alpha <= 0 || lambda <= 0) return(Inf)  # Ensure valid parameter range
  # Negative log-likelihood (using dgamma from extraDistr package)
  -sum(dgamma(data, shape = alpha, rate = lambda, log = TRUE))  # Negative log-likelihood
}
likelihood.gamma(c(2,1),data2$x)
likelihood.gamma(c(13.693, 0.711),data2$x)

# Initial guesses for parameters (alpha, lambda)
alpha.mle <- 2   # Example initial guess for alpha
lambda.mle <- 1  # Example initial guess for lambda
init <- c(alpha.mle, lambda.mle)
# Perform optimization
mle.gamma <- optim(
  par = init,
  fn = likelihood.gamma,
  data = data2$x,
  method = "L-BFGS-B",
  lower = c(0.1, 1e-6),  # Lower bounds for alpha and lambda
  upper = c(Inf, Inf),   # Upper bounds for alpha and lambda
  hessian = TRUE         # Compute Hessian matrix for parameter uncertainty
)

# Extract results
mle.gamma
det(mle.gamma$hessian)
```

## Conclusion

| Distribution | Dataset1 | Dataset2 |
|--------------|----------|----------|
| Normal       | ($\hat{\mu}$, $\hat{\sigma^2}$)= (0.025, 0.707) | ($\hat{\mu}$, $\hat{\sigma^2}$)= (19.247, 25.913)         |
| Exponential  | $\hat{\lambda}$=X                               | $\hat{\lambda}$=0.052         |
| Laplace      | ($\hat{\mu}$, $\hat{b^2}$)=(0.054,0.595)                               | ($\hat{\mu}$, $\hat{b^2}$)=(18.631,4.150)         |
| Gamma        | ($\hat{\alpha}$,$\hat{\lambda}$)=X                             | ($\hat{\alpha}$,$\hat{\lambda}$)=(13.693,0.711)         |

: MLE to estimate the model parameters

# (1.c) Are the MMEs and MLEs the same for each distribution?

觀察以下2個table，發現在normal和exponential在不同estimation是一樣的。

| Distribution | Dataset1 | Dataset2 |
|--------------|----------|----------|
| Normal       | ($\hat{\mu}$, $\hat{\sigma^2}$)= (0.025, 0.707) | ($\hat{\mu}$, $\hat{\sigma^2}$)= (19.247, 25.913)         |
| Exponential  | $\hat{\lambda}$=40.468                          | $\hat{\lambda}$=0.052        |
| Laplace      | ($\hat{\mu}$, $\hat{b^2}$)=(0.025,0.594)                               | ($\hat{\mu}$, $\hat{b^2}$)=(19.247,3.599)         |
| Gamma        | ($\hat{\alpha}$,$\hat{\lambda}$)=(0.00086,0.03496)                               | ($\hat{\alpha}$,$\hat{\lambda}$)=(14.296,0.743)         |

: MME to estimate the model parameters

| Distribution | Dataset1 | Dataset2 |
|--------------|----------|----------|
| Normal       | ($\hat{\mu}$, $\hat{\sigma^2}$)= (0.025, 0.707) | ($\hat{\mu}$, $\hat{\sigma^2}$)= (19.247, 25.913)         |
| Exponential  | $\hat{\lambda}$=X                               | $\hat{\lambda}$=0.052         |
| Laplace      | ($\hat{\mu}$, $\hat{b^2}$)=(0.054,0.595)                               | ($\hat{\mu}$, $\hat{b^2}$)=(18.631,4.150)         |
| Gamma        | ($\hat{\alpha}$,$\hat{\lambda}$)=X                             | ($\hat{\alpha}$,$\hat{\lambda}$)=(13.693,0.711)         |

: MLE to estimate the model parameters

# (1.d) and (1.e)

- 使用MME來找適合Dataset1和Dataset2的分布。
- 從(a)小題的圖可以看到Dataset1應該比較適合normal 或 laplace 而 Dataset2應該比較適合normal 或 gamma。

\hyperlink{page.4}{plot for normal}
\hyperlink{page.9}{plot for laplace}
\hyperlink{page.13}{plot for gamma}

# (1.f)

使用goodness-of-fit來檢定適合分布

## dataset1

### normal
```{r}
### Goodness-of-fit ###
### Hypothesis Testing using theoretical cdf ###
# install.packages("goftest")
library("goftest")
```
```{r}
n<-nrow(data1)# 樣本數量
EX<-mean(data1$x)# 一次動差
EX2<-mean(data1$x^2)# 二次動差
mu.mme<-EX
sigma2.mme<-EX2-(EX)^2

### K-S test ###
ks.N <- ks.test(data1$x, "pnorm", mean = mu.mme, sd = sqrt(sigma2.mme))
### AD test ###
ad.N <- ad.test(data1$x, "pnorm", mean = mu.mme, sd = sqrt(sigma2.mme))
### CVM test ###
cvm.N <- cvm.test(data1$x, "pnorm", mean = mu.mme, sd = sqrt(sigma2.mme))
# Summary #
dist.1 <- data.frame(normal = c(ks.N$p.value, ad.N$p.value, cvm.N$p.value))
```

### laplace

```{r}
library(extraDistr)
n <- nrow(data1) # 樣本數量
EX <- mean(data1$x) # 一次動差
EX2 <- mean(data1$x^2) # 二次動差
mu.mme <- EX
b.mme <- sqrt(0.5*(EX2-EX^2))

### K-S test ###
ks.L <- ks.test(data1$x, "plaplace", mu = mu.mme, sigma = b.mme)
### AD test ###
ad.L <- ad.test(data1$x, "plaplace", mu = mu.mme, sigma = b.mme)
### CVM test ###
cvm.L <- cvm.test(data1$x, "plaplace", mu = mu.mme, sigma = b.mme)
# Summary #
dist.1$laplace <- c(ks.L$p.value, ad.L$p.value, cvm.L$p.value)
kable(dist.1)
```

## dataset2

### normal

```{r}
n<-nrow(data2)# 樣本數量
EX<-mean(data2$x)# 一次動差
EX2<-mean(data2$x^2)# 二次動差
mu.mme<-EX
sigma2.mme<-EX2-(EX)^2

### K-S test ###
ks.N <- ks.test(data2$x, "pnorm", mean = mu.mme, sd = sqrt(sigma2.mme))
### AD test ###
ad.N <- ad.test(data2$x, "pnorm", mean = mu.mme, sd = sqrt(sigma2.mme))
### CVM test ###
cvm.N <- cvm.test(data2$x, "pnorm", mean = mu.mme, sd = sqrt(sigma2.mme))
# Summary #
dist.2 <- data.frame(normal = c(ks.N$p.value, ad.N$p.value, cvm.N$p.value))
```

### gamma

```{r}
n <- nrow(data2) # 樣本數量
EX <- mean(data2$x) # 一次動差
EX2 <- mean(data2$x^2) # 二次動差
alpha.mme <- EX^2/(EX2-(EX^2))
lambda.mme <- EX/(EX2-(EX^2))

### K-S test ###
ks.G <- ks.test(data2$x, "pgamma", shape = alpha.mme, rate = lambda.mme)
### AD test ###
ad.G <- ad.test(data2$x, "pgamma", shape = alpha.mme, rate = lambda.mme)
### CVM test ###
cvm.G <- cvm.test(data2$x, "pgamma", shape = alpha.mme, rate = lambda.mme)
# Summary #
dist.2$g <- c(ks.G$p.value, ad.G$p.value, cvm.G$p.value)
kable(dist.2)
```

# (1.g)

## dataset1

用MLE之結果算log-likelihood 之值
```{r}
# Load necessary libraries
library(MASS)    # For fitdistr
library(stats4)  # For AIC function if needed

n <- length(data1$x)
mu.mle <- mean(data1$x)
sig2.mle <- sum((data1$x - mu.mle)^2)/n

### Numerical method ###
likelihood.normal <- function(par, data){
  mu <- par[1]
  sig2 <- par[2]
  joint <- dnorm(data, mean = mu, sd = sqrt(sig2))
  #return(prod(joint))
  #return(sum(log(joint)))
  return(sum(log(joint)))
}

likelihood.n1 <- likelihood.normal(c(mu.mle, sig2.mle), data = data1$x) # loglikelihood
###########laplace
mu.mle <- median(data1$x)  # Median for mu
b.mle <- mean(abs(data1$x - mu.mle))  # Mean absolute deviation for b
### Numerical method ###
# Negative log-likelihood function for Laplace distribution
library(extraDistr)
likelihood.laplace <- function(par, data) {
  mu <- par[1]
  b <- par[2]
  if (b <= 0) return(Inf)  # Ensure valid parameter range
  sum(dlaplace(data, mu = mu, sigma = b, log = TRUE))  # Negative log-likelihood
}
likelihood.l1 <- likelihood.laplace(c(mu.mle, b.mle), data = data1$x) # loglikelihood

# Number of parameters
k_norm <- 2  # Normal: mean and sd
k_laplace <- 2  # Laplace: mean and scale

# Calculate AICs
AIC_norm <- -2 * likelihood.n1 + 2 * k_norm
AIC_laplace <- -2 * likelihood.l1 + 2 * k_laplace

# Compare results
results1 <- data.frame(
  Distribution = c("Normal", "Laplace"),
  LogLikelihood = c(likelihood.n1, likelihood.l1),
  AIC = c(AIC_norm, AIC_laplace)
)
kable(results1)
```

從結果來看，laplace表現的更好。

## dataset2

用MLE之結果算log-likelihood 之值
```{r}
# Load necessary libraries
library(MASS)    # For fitdistr
library(stats4)  # For AIC function if needed

n <- length(data2$x)
mu.mle <- mean(data2$x)
sig2.mle <- sum((data2$x - mu.mle)^2)/n

### Numerical method ###
likelihood.normal <- function(par, data){
  mu <- par[1]
  sig2 <- par[2]
  joint <- dnorm(data, mean = mu, sd = sqrt(sig2))
  #return(prod(joint))
  #return(sum(log(joint)))
  return(sum(log(joint)))
}

likelihood.n2 <- likelihood.normal(c(mu.mle, sig2.mle), data = data1$x) # loglikelihood
########### gamma
### Numerical method ###
# log-likelihood function for Gamma distribution (rate parameterization)
library(extraDistr)
likelihood.gamma <- function(par, data) {
  alpha <- par[1]
  lambda <- par[2]
  if (alpha <= 0 || lambda <= 0) return(Inf)  # Ensure valid parameter range
  # Negative log-likelihood (using dgamma from extraDistr package)
  sum(dgamma(data, shape = alpha, rate = lambda, log = TRUE))  #  log-likelihood
}
likelihood.g2 <- likelihood.gamma(c(13.6931953, 0.7114569), data = data2$x) # loglikelihood

k_norm <- 2
k_gamma <- 2
# Calculate AICs
AIC_norm <- -2 * likelihood.n2 + 2 * k_norm
AIC_gamma <- -2 * likelihood.g2 + 2 * k_gamma

# Compare results
results2 <- data.frame(
  Distribution = c("Normal", "Gamma"),
  LogLikelihood = c(likelihood.n2, likelihood.g2),
  AIC = c(AIC_norm, AIC_gamma)
)
kable(results2)
```

從此結果來看，gamma分布應該比較適合dataset2。

# iris data for question 2

# Load datasets iris.csv

```{r}
#| echo: true
#| code-fold: show
library(knitr)
iris <- read.csv("iris.csv") ; head(iris)
```

# EDA

```{r}
#| output: asis
#| code-fold: false
#| echo: false
library(Hmisc)
latex(describe(iris), file = "", caption.placement = "top")
```

# (2.a)

```{r}
n <- nrow(iris); n
petal.l <- iris[,3]
mu.mle <- mean(petal.l) ; mu.mle
sig2.mle <- sum((petal.l - mu.mle)^2)/n ; sig2.mle

### Numerical method ###
likelihood.normal <- function(par, data){
  mu <- par[1]
  sig2 <- par[2]
  joint <- dnorm(data, mean = mu, sd = sqrt(sig2))
  #return(prod(joint))
  #return(sum(log(joint)))
  return(-sum(log(joint)))
}
likelihood.normal(c(mu.mle, sig2.mle), data = petal.l) # -loglikelihood
likelihood.normal(c(3.757784, 3.095191), data = petal.l) # -loglikelihood

#optim(c(2, 1), likelihood.normal, data = dat.x, control=list(fnscale=-1))
mle.norm <- optim(c(2, 1), likelihood.normal, data = petal.l, hessian = TRUE)
mle.norm
```

By the MLE(Normal distribution), we can know that the parameters is:
$$
\hat{\mu}_{MLE} = 3.758
$$

$$
\hat{\sigma^2}_{MLE} = 3.095503
$$

# (2.b)

According to 2.a, we can know that $\hat{\mu}_{MLE}= 3.758$ by MLE method. This value is apparently greater than we assumed, 1.5.

# (2.c)

By 2.a, we know that the sampling distribution of $\hat{\mu}$ is $\hat{\mu} \sim N(3.758, \frac{3.095503}{150})$.

I decide to use the bootstrapping method to complete this process. The code is shown below:

```{r}
set.seed(24101222)
# 1. 定義參數
B <- 1000  # 自助抽樣次數
n <- length(petal.l)  # 樣本大小 150
# 2. 初始化用於存放均值的向量
bootstrap_means <- numeric(B)
# 3. 開始自助抽樣
for (i in 1:B) {
  # 隨機抽樣（有放回）
  resampled_data <- sample(petal.l, size = n, replace = TRUE)
  # 計算重抽樣後的均值
  bootstrap_means[i] <- mean(resampled_data)
}
# 4. 計算 95% 置信區間
ci_lower <- quantile(bootstrap_means, probs = 0.025)  # 下界（2.5% 分位數）
ci_upper <- quantile(bootstrap_means, probs = 0.975)  # 上界（97.5% 分位數）
# 5. 輸出結果
cat("95% Bootstrap Confidence Interval for Mean:\n")
cat("Lower Bound:", ci_lower, "\n")
cat("Upper Bound:", ci_upper, "\n")
```

According to the result shown above, the **95% confidence interval** of $\mu$ is $\mu \in [3.488, 4.032717]$

# (2.d)

By the result shown in 2.c, the statement which said that **the mean of petal length is 1.5** is not reasonable.

# (2.e)

$t \ statistic = \frac{\bar{x}- \mu}{s/\sqrt{n}}$ 

```{r}
# 1. 假設檢定參數
mu_0 <- 1.5  # 假設均值
alpha <- 0.05  # 顯著性水平

# 2. 計算 t 統計量
sample_mean <- mean(petal.l)  
sample_sd <- sd(petal.l)      
n <- length(petal.l)          
t_stat <- (sample_mean - mu_0) / (sample_sd / sqrt(n))  # t 統計量

# 3. 計算自由度與 p 值
df <- n - 1  # 自由度
p_value <- 2 * pt(-abs(t_stat), df)  # 雙尾檢定的 p 值

# 4. 檢查是否拒絕 H0
if (p_value < alpha) {
  cat("Reject H0: The mean is significantly different from 1.5.\n")
} else {
  cat("Fail to Reject H0: There is no significant evidence that the mean is different from 1.5.\n")
}

# 5. 顯示結果
cat("t-statistic:", t_stat, "\n")
cat("p-value:", p_value, "\n")
```

As the result shown in the code chunk, we reject the null hypothesis which said that the true mean is 1.5. This result have the same conclusion as the result in 2.d.